Species richness destabilizes trophic communities in freshwater ecosystems

### 2020-02-05 manuscript/bef_stability/manuscript.Rmd
Add numbers for the linear models

```{r get-data, cache = TRUE}
library(nlme)
library(piecewiseSEM)

@@

```{r mod-div-total}
library(lme4)
ctrl <- lmeControl(opt='optim')
ctrl_lmer <- lmerControl(optCtrl = list(method = 'optimx'))
stab_div_mod <- lmer(
  data = sem_data,
  control = ctrl_lmer,
  formula = log_stab ~ log_rich + (1 | basin)
)
stopifnot(!isSingular(stab_div_mod, tol = 1e-05))
#stab_div_colin <- car::vif(stab_div_mod) not use if one response
stab_div_sum <- piecewiseSEM::rsquared(stab_div_mod, method = "nagelkerke") %>%
  mutate_at(vars(Conditional, Marginal), ~round(., 2))

@@

#
bm_div_mod <- lmer(data = sem_data,
  formula = log_bm ~ log_rich + (1 | basin)
)
bm_div_sum <- piecewiseSEM::rsquared(bm_div_mod, method = "nagelkerke") %>%
  mutate_at(vars(Conditional, Marginal), ~round(., 2))

pred <- ggpredict(bm_div_mod, terms = c("log_rich"), type = "fe")
#plot(pred, add.data = TRUE)
bm_div_pred <- pred %>%
  as_tibble() %>%
  rename(
    log_bm = predicted,
    log_rich = x,
    basin = group
  )
```

```{r mod-div-troph, fig.show = "hide"}
stab_troph_mod <- lmer(data = na.omit(stab_div_troph),
  formula = log_stab ~ log_rich + troph_group + log_rich:troph_group + (1 | basin)
  #control = ctrl
)
mySumm2 <- function(.) {
  c(beta=fixef(.),sigma=sigma(.), sig01=sqrt(unlist(VarCorr(.))))
}
stab_troph_boot <- bootMer(stab_troph_mod, mySumm2, nsim = 100)
stab_troph_ci <- confint(stab_troph_boot)

@@

pred_troph_mod <- ggpredict(bm_troph_mod, terms = c("log_rich", "troph_group"), type = "fe")
#plot(pred_troph_mod, add.data = TRUE)
bm_troph_pred <- pred_troph_mod %>%
  as_tibble() %>%
  rename(
    log_bm = predicted,
    log_rich = x,
    troph_group = group
  )
```

@@

p_stab_div <- sem_data %>%
  ggplot(aes(x = log_rich, y = log_stab))+
  geom_point() +
  ylim(lim_log_stab) +
  xlim(lim_log_rich) +
  #scale_y_log10() +
  #scale_x_log10() +
  labs(x = "Median species richness (Log 10)", y = " Biomass stability (Log 10)") +
  geom_ribbon(data = stab_div_pred, aes(ymin = conf.low, ymax = conf.high), fill
  = "grey20", alpha = .2) +
  annotate("text", x = 1.0, y = .8, label = make_label_rsq(rsq = stab_div_sum),
    parse = TRUE)

@@

```{r psbm, fig.dim = c(10,5)}
plot_grid(p_stab, p_bm, ncol = 1)
```

- Stability:
  - decreases with richness
  - is higher for higher trophic group
  - increases and decreases with richness for the lower and higher trophic group
    respectively

- richness on stab:
  - globally negative
  -

@@

- Latex:
  - csv:
    - https://tex.stackexchange.com/questions/178904/use-datatool-to-read-a-row-from-a-csv-file-then-use-the-variables-in-the-docume
    - https://tex.stackexchange.com/questions/40901/datatool-getting-a-specific-value-given-the-value-of-another-column
  - path:

@@

```{r}
test <- map2_dfr(list(bm_troph_ci, stab_troph_ci), list(bm_troph_mod, stab_troph_mod),
  clean_summary_mod, term_rp = rp_est, group_rp = rp_group)

kable(test) %>%
  pack_rows("Biomass", 1, 6) %>%
  pack_rows("Stab", 7, 12)
```

@@

```{r}
```

>>>>

```{r get-data, cache = FALSE}
library(nlme)
library(piecewiseSEM)

@@

```{r mod-div-total}
library(lme4)
ctrl <- lmeControl(opt='optim')
ctrl_lmer <- lmerControl(optCtrl = list(method = 'optimx'))
stab_div_mod <- lmer(
  data = sem_data,
  control = ctrl_lmer,
  formula = log_stab ~ log_rich + (1 | basin)
)
stopifnot(!isSingular(stab_div_mod, tol = 1e-05))
stab_div_boot <- bootMer(stab_div_mod, mySumm2, nsim = 100)
stab_div_ci <- confint(stab_div_boot)
#stab_div_colin <- car::vif(stab_div_mod) not use if one response
stab_div_sum <- piecewiseSEM::rsquared(stab_div_mod, method = "nagelkerke") %>%
  mutate_at(vars(Conditional, Marginal), ~round(., 2))

@@

#
bm_div_mod <- lmer(data = sem_data,
  formula = log_bm ~ log_rich + (1 | basin)
)
bm_div_boot <- bootMer(bm_div_mod, mySumm2, nsim = 100)
bm_div_ci <- confint(bm_div_boot)
bm_div_sum <- piecewiseSEM::rsquared(bm_div_mod, method = "nagelkerke") %>%
  mutate_at(vars(Conditional, Marginal), ~round(., 2))

pred <- ggpredict(bm_div_mod, terms = c("log_rich"), type = "fe")
#plot(pred, add.data = TRUE)
bm_div_pred <- pred %>%
  as_tibble() %>%
  rename(
    log_bm = predicted,
    log_rich = x,
    basin = group
  )

# get coeff
coef_div <- map(list(stab = stab_div_mod, bm = bm_div_mod), fixed.effects)
```

```{r mod-div-troph, fig.show = "hide"}
stab_troph_mod <- lmer(data = na.omit(stab_div_troph),
  formula = log_stab ~ log_rich + troph_group + log_rich:troph_group + (1 | basin)
  #control = ctrl
)
stab_troph_boot <- bootMer(stab_troph_mod, mySumm2, nsim = 100)
stab_troph_ci <- confint(stab_troph_boot)

@@

pred_troph_mod <- ggpredict(bm_troph_mod, terms = c("log_rich", "troph_group"), type = "fe")
#plot(pred_troph_mod, add.data = TRUE)
bm_troph_pred <- pred_troph_mod %>%
  as_tibble() %>%
  rename(
    log_bm = predicted,
    log_rich = x,
    troph_group = group
  )

# get coeff
coef_troph <- map(list(stab = stab_troph_mod, bm = bm_troph_mod), fixed.effects)
```

@@

p_stab_div <- sem_data %>%
  ggplot(aes(x = log_rich, y = log_stab))+
  geom_point() +
  ylim(lim_log_stab) +
  xlim(lim_log_rich) +
  #scale_y_log10() +
  #scale_x_log10() +
  labs(x = "Median species richness (Log 10)", y = " Biomass stability (Log 10)") +
  geom_line(data = stab_div_pred) +
  geom_ribbon(data = stab_div_pred, aes(ymin = conf.low, ymax = conf.high), fill
  = "grey20", alpha = .2) +
  annotate("text", x = 1.0, y = .8, label = make_label_rsq(rsq = stab_div_sum),
    parse = TRUE)

@@

```{r psbm, fig.dim = c(10,5)}
plot_grid(p_stab, p_bm, ncol = 1)

print_coef <- function (vec){
  round(abs(vec)*100)
}
print_ci <- function(.df){
  .df <- round(.df, 3)*100

  paste0("[", .df[["2.5 %"]],";",.df[["97.5 %"]] ,"]")
}

```{r print-coef-stab}
# Coefficients
s_co_i <- print_coef(coef_div$stab["(Intercept)"])
s_co_r <- print_coef(coef_div$stab["log_rich"])
s_co_t <- print_coef(coef_troph$stab["troph_group3"])
s_co_r_t2 <- print_coef(coef_troph$stab["log_rich"])
s_co_r_t3 <- print_coef(coef_troph$stab["log_rich"]+coef_troph$stab["log_rich:troph_group3"])

s_ci_r <- print_ci(stab_div_ci["beta.log_rich",])
s_ci_t <- print_ci(stab_troph_ci["beta.troph_group3",])
s_ci_r_t2 <- print_ci(stab_troph_ci["beta.log_rich",])
s_ci_r_t3 <- print_ci(coef_troph$stab["log_rich"] + stab_troph_ci["beta.log_rich:troph_group3",])
```

```{r print-coef-bm}
# Coefficients
b_co_i <- print_coef(coef_div$bm["(Intercept)"])
b_co_r <- print_coef(coef_div$bm["log_rich"])
b_co_t <- print_coef(coef_troph$bm["troph_group3"])
b_co_r_t2 <- print_coef(coef_troph$bm["log_rich"])
b_co_r_t3 <- print_coef(coef_troph$bm["log_rich"]+coef_troph$bm["log_rich:troph_group3"])

b_ci_r <- print_ci(bm_div_ci["beta.log_rich",])
b_ci_t <- print_ci(bm_troph_ci["beta.troph_group3",])
b_ci_r_t2 <- print_ci(bm_troph_ci["beta.log_rich",])
b_ci_r_t3 <- print_ci(coef_troph$bm["log_rich"] + bm_troph_ci["beta.log_rich:troph_group3",])
```

- Stability:
  - decreases globally with median species richness (MSR)
    - An increase of 1% of MSR decreased biomass stability by
      `r s_co_r`% (`r s_ci_r`)
  - was `r s_co_t`% (`r s_ci_t`) higher for higher trophic group
  - increased by `r s_co_r_t2`% (`r s_ci_r_t2`) and decreased by `r s_co_r_t3`
    (`r s_ci_r_t3`) with 1% increase in MSR for the lower and higher trophic
    group respectively

- Biomass:
  - increases with MSR 
    - An increase of 1% of MSR increased total biomass by
      `r b_co_r` % `r b_ci_r`
  - was `r b_co_t`% `r b_ci_t` higher for higher trophic group
  - increased by `r b_co_r_t2`% `r b_ci_r_t2` and decreased by `r b_co_r_t3`
    (`r b_ci_r_t3`) with 1% increase in MSR for the lower and higher trophic
    group respectively

@@

- Latex:
  - csv:
    - https://tex.stackexchange.com/questions/178904/use-datatool-to-read-a-row-from-a-csv-file-then-use-the-variables-in-the-docume
    - https://tex.stackexchange.com/questions/40901/datatool-getting-a-specific-value-given-the-value-of-another-column
  - path: 
- Reporting:
  - equations: https://stats.stackexchange.com/questions/169115/reporting-the-actual-formula-equation-of-an-lme-model-with-factors-used-in-r
  - log-log results:
    - https://dev.to/rokaandy/logarithmic-transformation-in-linear-regression-models-why-when-3a7c
    - https://stats.stackexchange.com/questions/18480/interpretation-of-log-transformed-predictor-and-or-response

@@

```{r}
coeff_troph <- map2_dfr(list(bm_troph_ci, stab_troph_ci), list(bm_troph_mod, stab_troph_mod),
  clean_summary_mod, term_rp = rp_est, group_rp = rp_group)

kable(coeff_troph) %>%
  pack_rows("Biomass", 1, 6) %>%
  pack_rows("Stab", 7, 12)
```

@@

```{r, eval = FALSE}
compute_stab_sem_rich(.data = sem_data, get_sem =TRUE)

mod_list <-  
  list(
    nlme::lme(log_rich ~ RC1 + RC2 + RC3 + RC4 + RC5,  random = ~ 1 | basin, data = sem_data),
    nlme::lme(piel ~ RC1 + RC2 + RC3 + RC4 + RC5, random = ~ 1 | basin, data = sem_data),
    nlme::lme(ct ~ RC1 + RC2 + RC3 + RC4 + RC5 + log_rich, random = ~ 1 | basin, data = sem_data),
    nlme::lme(t_lvl ~ RC1 + RC2 + RC3 + RC4 + RC5 + log_rich, random = ~ 1 | basin, data = sem_data),
    #nlme::lme( ~ RC1 + RC2 + RC3 + RC4 + RC5 + log_rich, random = ~ 1 | basin, data = sem_data),
    nlme::lme(log_sync ~ log_rich + piel + ct + t_lvl + RC1 + RC2 + RC3 + RC4 + RC5,
      random = ~ 1 | basin, data = sem_data),
    nlme::lme(log_cv_sp ~ log_rich + piel + ct + t_lvl
      + RC1 + RC2 + RC3 + RC4 + RC5, random = ~ 1 | basin, data = sem_data),
    lm(log_stab ~ log_cv_sp + log_sync, data = sem_data)
  )
names(mod_list) <- c("richness", "pielou", "connectance", "weighted_mean_trophic_lvl", "synchrony", "cv_sp", "stability") 
colin_stab_sem <- map(mod_list, car::vif)

tibble(
  vif = colin_stab_sem,
  mod = names(mod_list)
) 
test <- cbind(colin_stab_sem)

formulas <- list(
  "log_rich ~ RC1 + RC2 + RC3 + RC4 + RC5",
  "piel ~ RC1 + RC2 + RC3 + RC4 + RC5"
)
test <- map(formulas, function(x, y, random_effect) {
  lmer(as.formula(x), random= as.formula(random_effect), data = y)

}, y =sem_data, random_effect = "~ 1 |basin")
